---
title: "Reg511HW1"
author: "Ben Straub"
date: "August 26, 2015"
output: html_document
---

1. For the following matrices, find the desired quantities. You may either do this by hand or using R. See
the ”IntroToR.r” code for some examples.
install.packages("matrixcalc")
```{r, echo=TRUE, results='hide'}
library(matrixcalc)

HI BEN LEE !!!!

A = matrix(
  c(1, 3, -1, 9, 5, -2, -1, -1, 2, 2, -3, 1),
  nrow = 4,
  ncol = 3)
A

b = matrix(
  c(2, 1, -1),
  nrow = 3,
  ncol = 1
)
b


```

(a) A*b
```{r, echo=TRUE}
A%*%b
```

(b) b`b
```{r, echo=TRUE}
bT<- t(b)
bT%*%b
```

(c) bb`
```{r, echo=TRUE}
bT<- t(b)
b%*%bT
```

(d) A`A
```{r, echo=TRUE}
AT<- t(A)
AT%*%A
```

(e) trace(A`A)
```{r, echo=TRUE}
AT<- t(A)
e <- AT%*%A
matrix.trace(e)
```

(f) (A`A)^-1
```{r, echo=TRUE}
AT<- t(A)
e <- AT%*%A
solve(e)
```


##########      6   ##################
```{r, echo=TRUE}
set.seed(6)
x <- rnorm(100, mean = 1, sd = 2)
y <- rnorm(100, mean = -3, sd = 1)

mean(x)
mean(y)

var(x)
var(y)

z <- x - y

mean(z)
var(z)
```
(a)  *The sample mean for x is 0.9795865, which is within 0.02 of the theoretical mean, 1.    
     *The sample mean for y is -3.08893, which is within 0.09 of the   
     theoretical mean, -3.  
     *The sample variance for x is 4.272303, which is within 0.27 of the 
     theoretical variance, 4.
     *The sample variance for x is 4.272303, which is within 0.27 of the     
     theoretical variance, 4.
     
(b)  *The sample mean for z is 4.068516, which is within .069 of the     
     theoretical mean, 4.
     *The sample variance for z is 4.387706, which is 0.612 of the theoretical      mean, 5