---
title: "STAT511HW5"
author: "Ben Straub"
date: "October 27, 2015"
output: pdf_document
---
```{r}
setwd("/Users/benStraub/Desktop/STAT511")
rm(list = ls())
smoking = read.csv("smoking.csv", sep=",")

library(knitr)
str(smoking)
attach(smoking)

pairs(smoking)
kable(summary(smoking), digits=2)

fit = lm(deaths~agecat+smokes+pyears, data=smoking)
summary(fit)
fit = lm(deaths~factor(agecat)+factor(smokes)+pyears, data=smoking)
summary(fit)
fit = lm(deaths~factor(agecat)+smokes+pyears, data=smoking)
summary(fit)
fit = lm(deaths~agecat+factor(smokes)+pyears, data=smoking)
summary(fit)


fit=glm(deaths~factor(agecat)+factor(smokes),offset=log(pyears),family="poisson",data=smoking)
summary(fit)

plot(fit)

library(car)
crPlots(fit)

## Test for outliers
library(car)
outlierTest(fit)


##
## simulation study to see if resids are acceptable
##

## simulate data from the fitted model
muhat=predict(fit,type="response")

ysim=rpois(length(muhat),lambda=muhat)

## fit the simulated data
fit.sim=glm(ysim~factor(agecat)+factor(smokes),offset=log(pyears),family="poisson",data=smoking)
summary(fit.sim)

## residual plots
plot(fit.sim)


###############################################
##
## Predictions 
##
###############################################

## predicted mean # of resp.deaths per year for 1000 people in a particular group
##  by setting "arsenic=1:4" I am jointly calculating everything for all 4 arsenic groups
##  This allows us to compare predicted respitory death rates across arsenic exposure levels

## make a data frame with the desired predictor variables
data=data.frame(pyears=1000,agegr=2,period=1,arsenic=1:4,start=1)
data

## get the mean and sd of the linear predictor eta for those predictor variables
pred.mean=predict(fit,newdata=data,type="response",se=T)
mu.hat=pred.mean$fit

## CI bounds on linear predictor
CI.up=mu.hat+1.96*pred.mean$se.fit
CI.down=mu.hat-1.96*pred.mean$se.fit

## CI on LINEAR PREDICTOR in table form
cbind(CI.down,CI.up)

## CI on MEAN # of resp.deaqths per year for 1000 people
## Note: for Poisson regression, response function is mu=exp(eta)
exp(cbind(CI.down,CI.up))
```


```{r}
setwd("/Users/benStraub/Desktop/STAT511")
rm(list = ls())
load("isj.Rdata")
names(isj)[names(isj)=="isj"] <- "birds"
library(knitr)
str(isj)
attach(isj)
pairs(isj)
kable(summary(isj), digits=2)
```

## Exploratory Data Analysis

* The Island Scrub Jay Data Set has 5,265 observations and 6 Variables
* I renamed the Variable isj as birds.
* Variables: birds, x, y, elev, forest, chap
* The birds Variable is a coded as 0 for abscence of birds and 1 for presence of birds.
* There appears to be a lot of NAs in the data set!!

```{r}
head(isj)
tail(isj, n= 10)
# Subsetting data on presence of birds
isj.sub_1 <- subset(isj, birds == 1)
# Subsetting data on absence of birds
isj.sub_0 <- subset(isj, birds == 0)
isj.sub_0_1 <- isj[complete.cases(isj),]
```

* Upon further examination of the data set I found that there was only 303 complete cases of Data.
* The Data has 38 entries for the presences of birds
* The Data has 265 entries for the absence of birds.


```{r}
fit=lm(birds~x+y+elev+forest+chap, data=isj.sub_1)
kable(summary(fit)$coef, digits=15)

fit=lm(birds~x+y+elev+forest+chap, data=isj.sub_1)
kable(summary(fit)$coef, digits=15)

fit=glm(birds~.,family=binomial(link="probit"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~.,family=binomial(link="logit"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~.,family=poisson(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~.,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~.,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~elev+forest+chap,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~forest+chap,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~forest,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~chap,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
fit=glm(birds~chap+x+y,family=binomial(link="log"),data=isj.sub_0_1)
summary(fit)
```

##  Creates the Map

```{r, echo=FALSE}
plot(isj[,2:3],type="n",main="Elevation",asp=1)
na.idx=which(is.na(isj$elev))
elev.star=isj$elev[-na.idx]
norm.elev=(elev.star-min(elev.star))/(max(elev.star)-min(elev.star))
points(isj[-na.idx,2:3],pch=20,col=grey(1-norm.elev))
points(isj[1:307,2:3],pch=4,col="red")
```

